Results exploration

### 2023-03-17 README.Rmd
updated figures, analysis, and readme based on Robbie's comments

<!-- badges: start -->
<!-- badges: end -->

The repository contains the data and the analysis for the research paper titled "A survey of registration practices among observational researchers using preexisting datasets". Other materials belonging to the project can be found on the project's OSF page at [link](link).

@@

>>>>

The repository contains the data and the analysis for the research paper titled "A survey of registration practices among observational researchers using preexisting datasets". Other materials belonging to the project can be found on the project's OSF page at [https://osf.io/x2dpn/](https://osf.io/x2dpn/).

@@

The `data/` folder contains all the datafiles and the corresponding codebooks that are needed to reproduce the results of the project.

The `preprocessing/` folder contains all the data preprocessing files in rmarkdowns. Within this folder you can find the following files:

  * `ecaw_source_raw_preprocessing.Rmd` file contains the code necessary for the transformation of the source data (the datafile downloaded directly from Qualtrics as is) to the raw datafile (datafile with standard naming). We also made sure that the raw datafile does not contain any information that can be used to indetify any of the respondents.
  * `ecaw_raw_processed_preprocessing.Rmd` file contains the code that cleans the dataset and transforms is into tidy format ready for the analysis.

The `manuscript/` folder contains the manuscript and supplementary materials documents with the analysis code embedded in them.

The `R/` folder contains any custom R functions and their documentation.

### 2023-03-17 manuscript/interactive_report.Rmd
updated figures, analysis, and readme based on Robbie's comments

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
library(likert)
library(viridis)
library(shiny)
# r_refs("references.bib")

@@

```{r, echo=FALSE}
inputPanel(
  sliderInput("n_studies", label = "Number of studies using ALSPAC:",
               min = min(processed$n_studies, na.rm = T), max = max(processed$n_studies, na.rm = T), value = max(processed$n_studies, na.rm = T), step = 1),
  radioButtons("methods", label = "Used the following methods:", choices = c("Blind", "Script", "Confirmatory", "Exploratory"), selected = NULL, inline = FALSE),
    radioButtons("concerned", label = "Concerned with rigour and reproducibility:", choices = c("Negative", "Neutral", "Support"), selected = NULL, inline = FALSE)
)

# Filter data
filtered_data <- reactive({
  processed %>% 
    filter(
      n_studies == input$n_studies
    )
  })
```

# Reproducibility and trustworthiness

The reproducible and trustworthiness ratings with the current ALSPAC system.

```{r}
# Prepare plot data
typically_plot_data <- 
  processed %>%
  select(contains("typically_")) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"))
    )) %>% 
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

# Create likert package data not including the missing values
typically_plot_data <- likert(typically_plot_data)

# Create figure
typically_plot <- 
  plot(typically_plot_data)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("papaja")
library(tidyverse)
library(here)
library(kableExtra)
library(gt)
library(patchwork)
library(likert)
library(viridis)
library(shiny)
library(plotly)
# r_refs("references.bib")

@@

```{r, echo=FALSE}
inputPanel(
  sliderInput("n_studies", label = "Number of studies using ALSPAC:",
              min = min(processed$n_studies, na.rm = T),
              max = max(processed$n_studies, na.rm = T), value = c(0, max(processed$n_studies, na.rm = T)), step = 1),
  selectInput("methods", label = "Used the following methods:",
              choices = c(
                "Blind" = "method_blind",
                "Script" = "method_script",
                "Preregistration" = "method_preregistered",
                "Confirmatory" = "method_confirmatory", 
                "Exploratory" = "method_exploratory"),
              selected = NULL, multiple = TRUE),
  selectInput("concerned", label = "Level of concern with rigour and reproducibility:",
              choices = c(
                "Less concerned",
                "Neutral",
                "More concerned"),
              selected = c(
                "Less concerned",
                "Neutral",
                "More concerned"), multiple = TRUE),
  selectInput("language", label = "Programming language used for the analysis:",
              choices = c(
                "R",
                "Stata",
                "SPSS",
                "SAS",
                "Python",
                "Mplus",
                "Bash",
                "MATLAB",
                "Nextflow",
                "plink2",
                "irrelevant"
              ),
              selected = c(
                "R",
                "Stata",
                "SPSS",
                "SAS",
                "Python",
                "Mplus",
                "Bash",
                "MATLAB",
                "Nextflow",
                "plink2",
                "irrelevant"
              ), multiple = TRUE)
  )

# Filter data
filtered_data <- reactive({
  concerned_options <-list(
    "Less concerned" =  c("very much less concerned", "less concerned", "somewhat less concerned"),
    "Neutral" =  c("as concerned as a typical researcher in my field"),
    "More concerned" = c("somewhat more concerned", "more concerned", "Very much more concerned")
  )

  concerned_filter <- unlist(concerned_options[input$concerned], use.names = FALSE)

  processed %>% 
    filter(n_studies >= input$n_studies[1] & n_studies <= input$n_studies[2] | is.na(n_studies)) %>% 
    filter(if_all(input$methods, ~ . != "Never or almost never")) %>% 
    filter(concerned %in% concerned_filter | is.na(concerned)) %>% 
    filter(str_detect(programming_language, paste(input$language, collapse="|")) | is.na(programming_language))
  })
# select(filtered_data(), concerned)
# shiny::renderTable(filtered_data())
shiny::renderText({nrow(filtered_data())})
```

```{r echo=FALSE, warning=FALSE}
# Create typically trustworthy and reproducible plot
# Prepare plot data
typically_plot_data <- reactive({
  filtered_data() %>%
    select(contains("typically_")) %>%
    mutate(
    across(
      .fns = ~ factor(., levels =   c(
    "Strongly disagree",
    "Somewhat disagree",
    "Neither agree nor disagree",
    "Somewhat agree",
    "Strongly agree"))
    )) %>%
    rename(
      Reproducible = typically_reproducible,
      Trustworthy = typically_trustworthy
      ) %>%
    # Create likert package data not including the missing values
    likert(.)
})

# Create figure
typically_plot <- reactive({
  plot(typically_plot_data(), digits = 1, ordered = FALSE, legend.position = "right") +
  labs(title = "Typically, studies that analyze preexisting observational datasets (such as the ALSPAC dataset) are...")
})

# Create ecaw trustworthy and reproducible plot
# Prepare plot data
ecaw_plot_data <- reactive({
  filtered_data() %>%
  select(contains("ecaw_")) %>%
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . == "I don't understand the question" ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>%
  mutate(
    across(
      .fns = ~ factor(., levels =   c(
          "Much less",
          "Somewhat less",
          "About the same",
          "Somewhat more",
          "Much more"))
    )) %>%
  rename(
    Reproducible = ecaw_reproducible,
    Trustworthy = ecaw_trustworthy
  ) %>%
    # Create likert package data not including the missing values
    likert(.)
})

# Create figure
ecaw_plot <- reactive({
  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right") +
  labs(title = "Compared to a typical study using preexisting observational data, a study using an ECAW would be...")
})

# Join the two plots together to make one figure
renderPlot({ typically_plot + ecaw_plot + plot_layout(ncol = 1, nrow = 2) })
```

```{r}
renderPlot({
# Create histogram of the number of observational studies done
filtered_data() %>% 
  # Exclude missing responses
  filter(!is.na(n_studies)) %>% 
  ggplot() +
  aes(x = n_studies) +
  geom_histogram(binwidth = 3) +
  # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  labs(
    x = "Number of studies",
    y = "Count",
    title = "Approximately how many studies have you published using a preexisting observational dataset (e.g., the ALSPAC dataset)?"
  ) +
  papaja::theme_apa()
})
```
